Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Mechanical Perturbations

### 2024-08-08 main.tex
Update new balance assist data with the values calculated without truncating the oscillation measurements.

In the early years of developments in automatic control, \citep{Whipple1899}
not only derived the correct equations of motion of the bicycle but realized
and showed that roll motion feedback can stabilize a bicycle. Much later,
attempts at automatic roll stabilization of single track vehicles began in
earnest after predictive motorcycle models were developed and refined in
throughout the 1970s. \Citet{vanZytveld1975} was influenced by Whipple's work
and seems to be the first to attempt to robotically stabilize a small motorbike
with a controlled inverted pendulum that mimicked rider lean, but he was not
successful in demonstrating what his theoretic control model correctly
predicted. In his model, he recognized that feedback of the vehicle roll angle
and angular rate was essential to stabilize the vehicle. It was not until the
early 1980s that \cite{Nagai1983} successfully demonstrated balancing a robotic
bicycle on a treadmill with both steering control and a laterally moving mass.
Ruijs and Pacjecka followed this by demonstrating an automatically balanced
motorcycle~\citep{Ruijs1986} and they did so with a steering motor. Ruijs and
Pacejka clearly showed that steer torque driven by roll angle feedback
stabilizes the capsize mode, roll angular rate feedback stabilizes the weave
mode, and steer angular rate feedback stabilizes the wobble mode. They also
proposed how the gains must change with respect to vehicle speed for favorable
control at all speeds. This roll motion feedback is the simplest controller
that can stabilize a single track vehicle above a minimum speed when one is not
concerned with wobble instabilities. Ruijs and Pacejka's work was not
particularly concerned with low speed stability and their vehicle was fully
automatic, no human rider was involved. Many more automatically balanced single
track vehicles have been demonstrated over the last 40 years, but none have
demonstrated that increasing low speed stability can assist in balancing and
possibly reduce single actor falls. Most of these robotic bicycle and
motorcycle designers did not intend for a human rider to also control the
stabilized vehicle. Nevertheless, an automatically stabilized bicycle can be
controlled by a human rider if the motor controlled steer torque and the rider
applied steer torque act on the steer in parallel.  The effect of this
automatic control gives the ability to change the dynamics, up to some limits,
of the human-controlled plant. In our prior study~\citep{Alizadehsaravi2023},
we demonstrate reduced motion during distractions due to the balance assist but
contrarily researchers with a similar vehicle recently showed rider
dissatisfaction with the stabilization~\cite{Hanakam2023}. \todo{Maybe this
paragraph can be a bit shorter? Of course important to mention previous
research, but all this info is not 100\% necessary for understanding our paper.
(I've been working on this paragraph for over a year for a few different
possible publications about the balance assist bike, I'd like to keep it in.
This is an attempt to set the record straight on where this entire idea of
doing "balance assist" derives from.-- JKM)}

>>>>

In the early years of developments in automatic control, \citep{Whipple1899}
not only derived the correct equations of motion of the bicycle but realized
and showed that roll motion feedback can stabilize a bicycle. Much later,
attempts at automatic roll stabilization of single track vehicles began in
earnest after predictive motorcycle models were developed and refined in
throughout the 1970s. \Citet{vanZytveld1975} was influenced by Whipple's work
and seems to be the first to attempt to robotically stabilize a small motorbike
with a controlled inverted pendulum that mimicked rider lean, but he was not
successful in demonstrating what his theoretic control model correctly
predicted. In his model, he recognized that feedback of the vehicle roll angle
and angular rate was essential to stabilize the vehicle. It was not until the
early 1980s that \cite{Nagai1983} successfully demonstrated balancing a robotic
bicycle on a treadmill with both steering control and a laterally moving mass.
Ruijs and Pacjecka followed this by demonstrating an automatically balanced
motorcycle~\citep{Ruijs1986} and they did so with a steering motor. Ruijs and
Pacejka clearly showed that steer torque driven by roll angle feedback
stabilizes the capsize mode, roll angular rate feedback stabilizes the weave
mode, and steer angular rate feedback stabilizes the wobble mode. They also
proposed how the gains must change with respect to vehicle speed for favorable
control at all speeds. This roll motion feedback is the simplest controller
that can stabilize a single track vehicle above a minimum speed when one is not
concerned with wobble instabilities. Ruijs and Pacejka's work was not
particularly concerned with low speed stability and their vehicle was fully
automatic, no human rider was involved. Many more automatically balanced single
track vehicles have been demonstrated over the last 40 years, but none have
demonstrated that increasing low speed stability can assist in balancing and
possibly reduce single actor falls. Most of these robotic bicycle and
motorcycle designers did not intend for a human rider to also control the
stabilized vehicle. Nevertheless, an automatically stabilized bicycle can be
controlled by a human rider if the motor controlled steer torque and the rider
applied steer torque act on the steer in parallel.  The effect of this
automatic control gives the ability to change the dynamics, up to some limits,
of the human-controlled plant. In our prior study~\citep{Alizadehsaravi2023},
we demonstrate reduced motion during distractions due to the balance assist but
contrarily researchers with a similar vehicle recently showed rider
dissatisfaction with the stabilization~\cite{Hanakam2023}.
